---
title: "Leaflet assignment"
author: "Naddao Thongsak"
date: "12 March 2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Leaflet assignment

In this report, we practice to view information in leaflet maps visualization. Leaflet is a library which can be integrated with R and be helpful to visualize data together with maps.
This report aim to use library leaflet with choropleth map to visualize population data in 77 provinces of Thailand(data in year 2024).

## Prepare population data
### Data of Thai population of each provinces in year 2024
Data of Thai population of each province in year 2024 was download from [Thai Official statistics registration systems](https://stat.bora.dopa.go.th/new_stat/webPage/statByYear.php).  

### Clean up data
Raw data has to be cleaned up, then selected only province name(in Thai) and population amount of each province.
```{r}
# Load raw data of population statistic of each province in Thailand
thai_province_population_2024_raw <- read.csv("stat_c67.txt", sep="|")
head(thai_province_population_2024_raw)

# filter only province name(in Thai), and population data
thai_province_population_2024 <- thai_province_population_2024_raw[, c(3,9)]
names(thai_province_population_2024) <- c("name_th", "population")

head(thai_province_population_2024)
```

### Calculation province population in percent of each province
```{r}
library(dplyr)

# calculate total population of Thai
total_population <- sum(thai_province_population_2024$population)

# calcuate percent population of each province in Thailand
thai_province_population_2024 <- thai_province_population_2024 %>%
                                    mutate(percent_pop = round( (population/total_population)*100, digit=2) )

```

## Load maps data of all provinces boundary in GeoJSON format
As choropleth map using boundary information in GeoJSON format, we load data of province boundary of each province in Thai for using later in map visualization.
### Load geojson map data
```{r}
library(jsonlite)
# Load GeoJSON of Thai province boundary data
thai_provinces_geojson <- sf::read_sf("simplified_provinces_geojson.geojson")

class(thai_provinces_geojson)
names(thai_provinces_geojson)
```

### Combine province population data to GeoJSON data
We combine Thai province population which we have been prepared with the boundary map in GeoJSON format
```{r}
# Combine population data to GeoJSON map data
thai_provinces_geojson_updated <- thai_provinces_geojson %>%   
                                      left_join(thai_province_population_2024, by= c("ADM1_TH"="name_th"))
head(thai_provinces_geojson_updated)
```

## Render maps data in leaflet using choropleth map
```{r}
library(leaflet)

# Define color bin using in choropleth map
bins <- c(0, 100000, 500000, 1000000, 2500000, 5000000, Inf)
bins <- as.numeric(bins)
pal <- colorBin("YlOrRd", domain = thai_provinces_geojson_updated$population, bins = bins)

# Define labels when mouse hover on province region
labels <- paste(thai_provinces_geojson_updated$ADM1_EN, ":", format( thai_provinces_geojson_updated$population, big.mark=",", scientific=FALSE), " people", "(", thai_provinces_geojson_updated$percent_pop,  "%)") 

# Render leaflet choropleth map
thai_provinces_geojson_updated %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolygons(
      fillColor = ~pal(population),
      weight = 2,
      opacity = 1,
      color = "white",
      dashArray = "3",
      fillOpacity = 0.7, 
      highlightOptions = highlightOptions(
        weight = 5,
        color = "#666",
        dashArray = "",
        fillOpacity = 0.7,
        bringToFront = TRUE),
      label = labels,
      labelOptions = labelOptions (
        style = list("font-weight" = "normal", padding = "3px 8px"),
        textsize = "15px",
        direction = "auto")) %>%
  addLegend(pal = pal, values = ~population, opacity = 0.7, title = NULL,
  position = "bottomright")
```
